# HDDM Example

HDDM_A uses Hoeffding bounds to test whether the mean of the monitored error stream has increased more than expected under stationarity. In practice, this makes it a detector for **real concept drift**.

Reference: Frias-Blanco, I., del Campo-Avila, J., Ramos-Jimenez, G., Morales-Bueno, R., Ortiz-Diaz, A., and Caballero-Mota, Y. (2015). *Online and nonparametric drift detection methods based on Hoeffding's bounds*. IEEE Transactions on Knowledge and Data Engineering, 27(3), 810-823. <doi:10.1109/TKDE.2014.2345382>

## Learning goal

This example demonstrates how to use a statistically grounded error-based detector within the standard Heimdall workflow.

```{r}
# Load Heimdall and the synthetic stream example.
library(heimdall)
```

```{r}
# Fix the seed for reproducibility.
seed <- 1
set.seed(seed)
```

```{r}
# Load the stream and derive the binary monitored signal.
data(st_drift_examples)
data <- st_drift_examples$univariate
data$prediction <- st_drift_examples$univariate$serie > 4
```

```{r}
# Plot the binary stream that HDDM_A will monitor.
plot(x=seq_len(nrow(data)), y=data$prediction)
```

```{r}
# Instantiate the HDDM_A detector.
model <- dfr_hddm()
```

```{r}
# Process the stream sequentially and record the detector alarms.
detection <- NULL
output <- list(obj=model, drift=FALSE)
for (i in seq_len(nrow(data))){
  output <- update_state(output$obj, data$prediction[i])
  if (output$drift){
    type <- 'drift'
    output$obj <- reset_state(output$obj)
  } else {
    type <- ''
  }
  detection <- rbind(detection, data.frame(idx=i, event=output$drift, type=type))
}
```

```{r}
# Print the detected drift points.
detection[detection$type == 'drift',]
```

```{r}
# Overlay the alarms on the original numeric stream.
plot(x=seq_len(nrow(data)), y=data$serie)
for (drift_index in detection[detection$type == 'drift', 'idx']) {
  abline(v=drift_index, col='red', lty=2)
}
```
