# EDDM Example

EDDM extends DDM by monitoring the distance between classification errors instead of only the error rate. This makes it especially useful when **real concept drift** is gradual rather than abrupt.

Reference: Baena-Garcia, M., del Campo-Avila, J., Fidalgo, R., Bifet, A., Gavaldà, R., and Morales-Bueno, R. (2006). *Early drift detection method*. Fourth International Workshop on Knowledge Discovery from Data Streams.

## Learning goal

This example is a practical template for users who want to monitor gradual degradation in predictive behavior.

```{r}
# Load Heimdall and the built-in synthetic stream.
library(heimdall)
```

```{r}
# Fix the seed so the results are reproducible.
seed <- 1
set.seed(seed)
```

```{r}
# Load the stream and derive the binary monitored signal.
data(st_drift_examples)
data <- st_drift_examples$univariate
data$prediction <- st_drift_examples$univariate$serie > 4
```

```{r}
# Visualize the binary stream that EDDM will monitor.
plot(x=seq_len(nrow(data)), y=data$prediction)
```

```{r}
# Instantiate the EDDM detector.
model <- dfr_eddm()
```

```{r}
# Process the stream sequentially and store each detected alarm.
detection <- NULL
output <- list(obj=model, drift=FALSE)
for (i in seq_len(nrow(data))){
  output <- update_state(output$obj, data$prediction[i])
  if (output$drift){
    type <- 'drift'
    output$obj <- reset_state(output$obj)
  } else {
    type <- ''
  }
  detection <- rbind(detection, data.frame(idx=i, event=output$drift, type=type))
}
```

```{r}
# Print the drift alarms returned by EDDM.
detection[detection$type == 'drift',]
```

```{r}
# Show those alarms on top of the original numeric series.
plot(x=seq_len(nrow(data)), y=data$serie)
for (drift_index in detection[detection$type == 'drift', 'idx']) {
  abline(v=drift_index, col='red', lty=2)
}
```
